---
title: "Getting started"
author: "Colin Millar"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

First lets create some data and fit a basic model, this will introduce the format of the data required and also the basics of model fitting.  The most basic sample is one in which there are 3 passes and 2 lifestages; this constitutes a single sample.

```{r}
# create a single electrofishing site visit with 3 passes and 2 lifestages
ef_data <- data.frame(n     = c(100, 53, 24, 50, 26, 12),
                   pass  = c(  1,  2,  3,  1,  2,  3),
                   stage = c(  1,  1,  1,  2,  2,  2))
ef_data
```

A sensible model to fit to this is one where the capture probability is constant accross passes but varies by lifestage.  This can be acheived by

```{r, eval=FALSE}
# Fit a simple model
m1 <- efp(n ~ 1 + factor(stage), data = ef_data, pass = pass)
```

Notice that it is nessicary to supply the information on which electrifishing pass via the argument 'pass'. 



Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
